---
title: "Cambridge retail vacancies"
author: ""
date: '`r Sys.Date()`'
output: kjutil::small_format
---

```{r parameters, echo=FALSE,include=FALSE,message=FALSE}
knitr::opts_chunk$set(echo=FALSE,fig.width=10, fig.height=8, comment=NA, warning=FALSE, message=FALSE)
```

# Retail vacancies

An exploration of vacant retail storefronts in Cambridge in April, 2020.


```{r setup}
library(tidyverse)
library(DataExplorer)
library(gt)
library(leaflet)
library(sf)
```

```{r data}
df_raw = read_csv(here::here('data/Retail_Vacant_Storefronts_April_2020.csv')) %>% 
  st_as_sf(wkt='Coordinates for Mapping', crs=4326)
df = df_raw %>% 
  mutate(`New Space`=`Former Tenant`=='New Space',
         `Length of Vacancy`=ordered(`Length of Vacancy`, 
            levels=c("Less than 6 months", "6-10 months", "1-2 years", "2 years or more")))
```

The Cambridge data set includes `r nrow(df)` properties in
`r n_distinct(df[['Commercial District']])` commercial districts.

## Distribution of categorical variables

```{r barchart}
df_explore = df %>% 
  st_drop_geometry() %>% 
  select(-(Address:`Zip Code`),
         -(`Former Tenant`:`Police Response Districts`), -(`Leasing Contact`:`Census Blocks 2010`))

plot_bar(df_explore, ggtheme=silgelib::theme_plex())
```

## Length of vacancy by district

```{r}
length_table = df %>% as.data.frame() %>% 
  count(`Commercial District`, `Length of Vacancy`, name='Count', sort=TRUE) %>% 
  pivot_wider(names_from=`Length of Vacancy`, 
              values_from=Count, values_fill=list(Count=0)) %>% 
  select(`Commercial District`, levels(df$`Length of Vacancy`))
length_table$Total=rowSums(length_table[-1])
length_table = length_table %>% 
  arrange(desc(Total))

gt(length_table) %>% 
  grand_summary_rows(columns=2:6, fns=list(Total=~round(sum(.), 0)))

# district_icons =  awesomeIcons(
#   icon = 'store-alt',
#   iconColor = 'black',
#   library = 'fa',
#   markerColor = getColor(df.20)
# )
# 
# leaflet(
#   #width='1000px', height='1000px',
#         options=leafletOptions(maxZoom=20)) %>% 
#   # addProviderTiles('CartoDB.Positron', group='Map') %>% 
#   # addProviderTiles('Esri.WorldImagery', group='Satellite') %>% 
#  addTiles("http://tiles.mapc.org/basemap/{z}/{x}/{y}.png",
#           attribution='Map tiles by <a href="http://mapc.org">MAPC</a>, Data by <a href="http://www.mass.gov/mgis/">MassGIS</a>.') %>% 
#   setView(-71.128184, 42.3769824, zoom = 14) %>% 
#   addAwesomeMarkers(data=df, group='District')
```

## Commercial districts

This map shows commercial districts as defined by Cambridge CDD.

```{r map}
dists = st_read(here::here('COD/ASSESSING_CommercialDistrictsFY2020.gdb.zip'),
                quiet=TRUE)
ggplot(dists) +
  geom_sf(fill='cadetblue2') +
  geom_sf_text(data=st_centroid(dists), aes(label=DIST_NAME), color='black',
               size=3) +
  guides(fill='none') +
  theme_void()
```

# Sources

- Vacancy data from [Cambridge Open Data](https://data.cambridgema.gov/Economic-Development/Retail-Vacant-Storefronts-April-2020/exmr-dxma)
- Map data from [Cambridge GIS](http://gis.cambridgema.gov)
